---
title: 'Practical Machine Learning: Course Project'
author: "Calvin Chin"
output: 
  html_document: 
    keep_md: yes
---

# Synopsis 

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. In this project, the author will use data from accelerometers to build a machine learning algorithm that will predict activity quality from activity monitors.

# Data

The **training** data for this project are available here: 
<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv>

The **test** data are available here: 
<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv>

NOTE: The data for this project come from this source: <http://groupware.les.inf.puc-rio.br/har>

# Preparing the environment

```{r, warning=FALSE}
# Load the required libraries
library(caret)
library(randomForest)

trainingRAW <- read.csv(file="pml-training.csv")
testingRAW <- read.csv(file="pml-testing.csv")

```


# Pre-processing

```{r, warning=FALSE}
# Remove first 7 columns that are not relevant to the model building
trainingNew <- trainingRAW[ , -c(1:7)]


# Remove columns that have too many NAs
isNA <- sapply(trainingNew, function(x) mean(is.na(x))) > 0.95

trainingNew <- trainingNew[ , isNA==FALSE]


# Remove columns with Near Zero Variance
indexNZV <- nearZeroVar(trainingNew)

trainingNew <- trainingNew[ , -indexNZV]

```


# Model Building

```{r, warning=FALSE}
# Create train and test data partitions
inTrain <- createDataPartition(y=trainingNew$classe, p=0.7, list=FALSE)

training <- trainingNew[inTrain,]
testing <- trainingNew[-inTrain,]


# Create Model Fit using Random Forest. Cross validation (i.e cv) is used as trainControl method, with number of resampling iterations set to 3 for faster performance
modFit <- train(classe ~ ., data=training, method="rf", trControl=trainControl(method="cv", number=3))

print(modFit$finalModel)
```


# Cross Validation
```{r, warning=FALSE}
pred_testing <- predict(modFit, newdata=testing)

confusionMatrix(pred_testing, testing$classe)
```
Cross Validation results show overall accuracy of the model is >99%


# Now, apply Model Fit to predict the 'classe' value for the test cases provided
```{r, warning=FALSE}
# For the Testing dataset, Extract only the same columns prepared from the Training dataset
requiredColumns <- names(trainingNew[,1:length(trainingNew)-1])
testingNew <- testingRAW[ , requiredColumns]


pred_testingNew <- predict(modFit, newdata=testingNew)

# First print prediction results onto the screen 
print(pred_testingNew)


# Next, write prediction results onto text files required for Course Project Part 2 Submission (1 charcter in each file)

for(i in 1:length(pred_testingNew))
{
    filename = paste0("problem_id_", i, ".txt")
    write.table(pred_testingNew[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
}

```


